---
title: "Analysis of Election Tweets"
author: "Andy Pickering"
date: "9/28/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Being in the middle (or I guess thankfully near the end!) of election season, I thought it would be interesting to examine tweets about the election.

Much of this analysis applies what I learned in a DataCamp course on text-mining: <https://www.datacamp.com/courses/intro-to-text-mining-bag-of-words>.

To collect the tweets, i'm using the rtweet package. The twitter API only lets you access tweets for the previous 7 days, so i've downloaded a collection of tweets spanning the first presidential debate and saved it as a csv file. 

The code used to download the tweets was:

```{r}
library(rtweet)

```

Next I kept only the tweets that were not retweets, and removed username mentions from the tweets:
```{r}

```

Next we need to clean up the tweets some more. I'm going to use the tm package for this.
```{r}
library(tm)

```
